Interval Probability and Measurement for Fuzzy Quantum Theories

### 2016-09-24 prob.tex
prob.tex Adding one more approach...

>>>>

\section{(The other possible approach keep Section \ref{sec:Quantum-Probability-Spaces}) }

\subsection{Question the original assumption(?)}

\section{(The other possible approach keep Section \ref{sec:Quantum-Probability-Spaces}) }

If we compare the counterexample in example~\ref{ex:three-dimensional-three-value}
and \ref{ex:three-dimensional-4-value}, we found example~\ref{ex:three-dimensional-4-value}
is not as bad as example~\ref{ex:three-dimensional-three-value}.
At least, the states map to $\imposs$ are all in the same plane in
example~\ref{ex:three-dimensional-4-value}. Maybe interval-valued
probability measures will be closer and closer to real-valued probability
measures if the intervals are more and more precise?

\subsection{Examples for Quantum Interval-valued Probability Measures}

The following example verify the core of a quantum interval-valued
probability measure is nonempty even if there is no natural correspondence.
For example, if $\mathscr{I}=\left\{ \imposs,\unlikely,\likely,\necess\right\} $,
there is no natural way to map from $\frac{1}{2}$, as both $\frac{1}{2}\in\left[0,\frac{1}{2}\right]=\unlikely$
and $\frac{1}{2}\in\left[\frac{1}{2},1\right]=\likely$. However,
we still can find a quantum interval-valued probability measure corresponding
to a quantum probability measure as in the following example.

\begin{example}[One-qubit quantum interval-valued probability measure]\label{ex:One-qubit-quantum-interval}
Given a two dimensional Hilbert space, a vector $\ket{\psi}$ can
be normalized and written in $\left(\begin{array}{c}
\cos\theta\\
\rme^{\rmi\gamma}\sin\theta
\end{array}\right)$, where $0\le\theta\le\frac{\pi}{2}$ and $0\le\gamma<2\pi$. Consider
a quantum interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$
defined by 
\begin{eqnarray*}
\bar{\mu}(\mathbb{0}) & = & \imposs\textrm{ ,}\\
\bar{\mu}(\mathbb{1}) & = & \necess\textrm{ ,}\\
\bar{\mu}(\proj{\psi}) & = & \begin{cases}
\necess & \textrm{, if }\theta=0\textrm{ ;}\\
\likely & \textrm{, if }0<\theta<\frac{\pi}{4}\textrm{ ;}\\
\likely & \textrm{, if }\theta=\frac{\pi}{4}\textrm{ and }0\le\gamma<\pi\textrm{ ;}\\
\unlikely & \textrm{, if }\theta=\frac{\pi}{4}\textrm{ and }\pi\le\gamma<2\pi\textrm{ ;}\\
\unlikely & \textrm{, if }\frac{\pi}{4}<\theta<\frac{\pi}{2}\textrm{ ;}\\
\imposs & \textrm{, if }\theta=\frac{\pi}{2}\textrm{ .}
\end{cases}
\end{eqnarray*}
Let $\ket{\psi^{\perp}}$ be the state perpendicular to $\ket{\psi}$.
Then, the vector representation of $\ket{\psi^{\perp}}$ can be simplified
as 
\[
\left(\begin{array}{c}
-\rme^{-\rmi\gamma}\sin\theta\\
\cos\theta
\end{array}\right)=-\rme^{-\rmi\gamma}\left(\begin{array}{c}
\cos\left(\frac{\pi}{2}-\theta\right)\\
\rme^{\rmi\left(\pi+\gamma\right)}\sin\left(\frac{\pi}{2}-\theta\right)
\end{array}\right)
\]
so we have 
\[
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi}\right)=\bar{\mu}\left(\proj{\psi^{\perp}}\right)=1\tensor*[_{\mathscr{I}}]{-}{}\bar{\mu}\left(\proj{\psi}\right)\textrm{ .}
\]
Hence, $\bar{\mu}$ is a quantum $\mathscr{I}$-interval-valued probability
measure. Notice that for all projection~$P$, $\mu_{\ket{0}}^{B}(P)\in\bar{\mu}(P)$
implies that $\bar{\mu}(P)$ is one of quantum $\mathscr{I}$-interval-valued
probability measures corresponding to $\mu_{\ket{0}}^{B}(P)$. However,
$\mu_{\ket{0}}^{B}(P)$ can correspond to infinity many quantum $\mathscr{I}$-interval-valued
probability measures by twisting the definition involving $\gamma$.\qed\end{example}

We are going to provide an quantum 4-value interval-valued probability
measures corresponding to no quantum probability measure, i.e., $\mu_{\rho}^{B}(P)\notin\bar{\mu}(P)$.
In general, we believe this kind of quantum $\mathscr{I}$-interval-valued
probability measure exists for any finite set of interval~$\mathscr{I}$.

\begin{example}[Three-dimensional quantum 4-value interval-valued
probability measure]\label{ex:three-dimensional-4-value} \footnote{\yutsung{Need to verify this example is convex!}}Given
a three dimensional Hilbert space with an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $.
Consider\emph{ }$\mathscr{I}=\left\{ \imposs,\unlikely,\likely,\necess\right\} $,
$\ket{\ps}=\frac{1}{\sqrt{2}}(\ket{0}+\ket{1})$, $\ket{\ms}=\frac{1}{\sqrt{2}}(\ket{0}-\ket{1})$,
and a quantum $\mathscr{I}$-interval-valued probability measure~$\bar{\mu}$
defined as follow: 
\begin{enumerate}
\item $\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs$
and $\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})=\bar{\mu}(\mathbb{1}-\proj{\ps})=\necess$. 
\item Consider the states orthogonal to $\ket{0}$ and $\ket{\ps}$ respectively,
$\ket{0_{\theta,\gamma}^{\perp}}=\rme^{\rmi\gamma}\sin\theta\ket{1}+\cos\theta\ket{2}$
and $\ket{\ps_{\theta,\gamma}^{\perp}}=-\rme^{\rmi\gamma}\sin\theta\ket{\ms}+\cos\theta\ket{2}$,
where $0\le\theta\le\frac{\pi}{2}$ and $0\le\gamma<2\pi$. Half of
those states need special treatment, i.e., when where $0\le\theta<\frac{\pi}{2}$
and $0\le\gamma<\pi$, we define $\bar{\mu}(\proj{0_{\theta,\gamma}^{\perp}})=\ket{\ps_{\theta,\gamma}^{\perp}}=\likely$
and $\bar{\mu}(\mathbb{1}-\proj{0_{\theta,\gamma}^{\perp}})=\bar{\mu}(\mathbb{1}-\proj{\ps_{\theta,\gamma}^{\perp}})=\unlikely$. 
\item Otherwise, $\bar{\mu}(\proj{\psi})=\unlikely$ and $\bar{\mu}(\mathbb{1}-\proj{\psi})=\likely$. 
\end{enumerate}
$\bar{\mu}$ cannot correspond to any quantum probability measure
because of the exact same reason as in example~\ref{ex:three-dimensional-three-value}.
Before we verify that $\bar{\mu}$ is a quantum interval-valued probability
measure, we need a notation for cross product. For $\ket{\psi}=\alpha_{0}\ket{0}+\alpha_{1}\ket{1}+\alpha_{2}\ket{2}=\begin{pmatrix}\alpha_{0}\\
\alpha_{1}\\
\alpha_{2}
\end{pmatrix}$ and $\ket{\phi}=\beta_{0}\ket{0}+\beta_{1}\ket{1}+\beta_{2}\ket{2}=\begin{pmatrix}\beta_{0}\\
\beta_{1}\\
\beta_{2}
\end{pmatrix}$, their cross product is defined as follow. 
\[
\ket{\psi\times\phi}=\ket{\psi}\times\ket{\phi}=\begin{vmatrix}\alpha_{1}^{*} & \beta_{1}^{*}\\
\alpha_{2}^{*} & \beta_{2}^{*}
\end{vmatrix}\ket{0}+\begin{vmatrix}\alpha_{2}^{*} & \beta_{2}^{*}\\
\alpha_{0}^{*} & \beta_{0}^{*}
\end{vmatrix}\ket{1}+\begin{vmatrix}\alpha_{0}^{*} & \beta_{0}^{*}\\
\alpha_{1}^{*} & \beta_{1}^{*}
\end{vmatrix}\ket{2}\textrm{ ,}
\]
where $\begin{vmatrix}\alpha & \beta\\
\alpha' & \beta'
\end{vmatrix}=\alpha\beta'-\alpha'\beta$. Notice that we put the complex conjugate in the cross product, so
that the result of cross product is orthogonal to $\ket{\psi}$ and
$\ket{\phi}$. 
\begin{eqnarray*}
\ip{\psi\times\phi}{\psi} & = & \left(\alpha_{1}\beta_{2}-\alpha_{2}\beta_{1}\right)\alpha_{0}+\left(\alpha_{2}\beta_{0}-\alpha_{0}\beta_{2}\right)\alpha_{1}+\left(\alpha_{0}\beta_{1}-\alpha_{1}\beta_{0}\right)\alpha_{2}=0\textrm{ ,}\\
\ip{\psi\times\phi}{\phi} & = & \left(\alpha_{1}\beta_{2}-\alpha_{2}\beta_{1}\right)\beta_{0}+\left(\alpha_{2}\beta_{0}-\alpha_{0}\beta_{2}\right)\beta_{1}+\left(\alpha_{0}\beta_{1}-\alpha_{1}\beta_{0}\right)\beta_{2}=0\textrm{ .}
\end{eqnarray*}
Also, even if $\ket{\psi}$ and $\ket{\phi}$ are normalized, their
cross product~$\ket{\psi\times\phi}$ need not be normalized as usual.

To verify $\bar{\mu}$ is a quantum interval-valued probability measure,
it is sufficient to check 
\begin{equation}
\begin{aligned}\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi_{1}}\right) & \subseteq\bar{\mu}\left(\proj{\psi_{2}}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(\proj{\psi_{3}}\right)\\
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi_{2}}\right) & \subseteq\bar{\mu}\left(\proj{\psi_{3}}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(\proj{\psi_{1}}\right)\\
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi_{3}}\right) & \subseteq\bar{\mu}\left(\proj{\psi_{1}}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(\proj{\psi_{2}}\right)
\end{aligned}
\label{eq:non-additive-vectors}
\end{equation}
for every orthonormal basis $\mathcal{B}=\left\{ \ket{\psi_{1}},\ket{\psi_{2}},\ket{\psi_{3}}\right\} $.
We are going to enumerate all possible orthonormal bases to verify
the equations, and case 1 and 2 are illustrated in figure~\ref{fig:three-dimensional-4-value}.
\begin{figure}
\includegraphics{\string"Three-dimensional quantum 4-value interval-valued probability measure\string".pdf}\caption{\label{fig:three-dimensional-4-value}This figure illustrates case
1 and 2 in $\mathbb{R}^{3}$ in example~\ref{ex:three-dimensional-4-value}.
The red and green dotted vector are $\ket{0}$ and $\ket{\ps}$ respectively.
All possible vectors of $\ket{0_{\theta,\gamma}^{\perp}}$ and $\ket{\ps_{\theta,\gamma}^{\perp}}$
are drawn in the red and green circles, respectively. Within the circles,
a vector~$\ket{\psi}$ in dotted part means $\bar{\mu}(\proj{\psi})=\likely$;
otherwise, $\bar{\mu}(\proj{\psi})=\unlikely$. The gray vector is
a generic vector~$\ket{0_{\theta,\gamma}^{\perp}}$, and the red
and green solid vectors are normalized $\ket{0}\times\ket{0_{\theta,\gamma}^{\perp}}$
and $\ket{\ps}\times\ket{0_{\theta,\gamma}^{\perp}}$, respectively.}
\end{figure}

\begin{enumerate}
\item When $\ket{\psi_{1}}$ is $\ket{0}$, then $\ket{\psi_{2}}$ and $\ket{\psi_{3}}$
must be $\ket{0_{\theta,\gamma}^{\perp}}$ for some $\theta$ and
$\gamma$. Because of 
\[
\ket{0}\times\ket{0_{\theta,\gamma}^{\perp}}=\begin{pmatrix}1\\
0\\
0
\end{pmatrix}\times\begin{pmatrix}0\\
\rme^{\rmi\gamma}\sin\theta\\
\cos\theta
\end{pmatrix}=\begin{pmatrix}0\\
-\cos\theta\\
\rme^{-\rmi\gamma}\sin\theta
\end{pmatrix}=\rme^{-\rmi\gamma}\begin{pmatrix}0\\
\rme^{\rmi\left(\pi+\gamma\right)}\sin\left(\frac{\pi}{2}-\theta\right)\\
\cos\left(\frac{\pi}{2}-\theta\right)
\end{pmatrix}=\rme^{-\rmi\gamma}\ket{0_{\frac{\pi}{2}-\theta,\pi+\gamma}^{\perp}}
\]
the pair $\ket{\psi_{2}}=\ket{0_{\theta,\gamma}^{\perp}}$ and $\ket{\psi_{3}}=\ket{0_{\frac{\pi}{2}-\theta,\pi+\gamma}^{\perp}}$
for ($0<\theta<\frac{\pi}{2}$ and $0\le\gamma<\pi$) or $\theta=0$
enumerate all possible orthonormal bases. Equation (\ref{eq:non-additive-vectors})
can be verified as follow. 
\begin{eqnarray*}
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{0}\right)=\necess & \subseteq & \likely\tensor*[_{\mathscr{I}}]{+}{}\unlikely=\bar{\mu}\left(\proj{0_{\theta,\gamma}^{\perp}}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(\proj{0_{\frac{\pi}{2}-\theta,\pi+\gamma}^{\perp}}\right)\\
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{0_{\theta,\gamma}^{\perp}}\right)=\unlikely & = & \unlikely\tensor*[_{\mathscr{I}}]{+}{}\imposs=\bar{\mu}\left(\proj{0_{\frac{\pi}{2}-\theta,\pi+\gamma}^{\perp}}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(\proj{0}\right)\\
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{0_{\frac{\pi}{2}-\theta,\pi+\gamma}^{\perp}}\right)=\likely & = & \imposs\tensor*[_{\mathscr{I}}]{+}{}\likely=\bar{\mu}\left(\proj{0}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(\proj{0_{\theta,\gamma}^{\perp}}\right)
\end{eqnarray*}
Similarly, when $\ket{\psi_{1}}$ is $\ket{\ps}$, equation (\ref{eq:non-additive-vectors})
holds. 
\item When $\ket{0}\notin\mathcal{B}$ and $\ket{\psi_{1}}=\ket{0_{\theta,\gamma}^{\perp}}$
for $0\le\theta<\frac{\pi}{2}$ and $0\le\gamma<\pi$, we want to
prove $\bar{\mu}\left(\proj{\psi_{2}}\right)=\bar{\mu}\left(\proj{\psi_{3}}\right)=\unlikely$.
Then, we have 
\begin{eqnarray*}
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{0_{\theta,\gamma}^{\perp}}\right)=\unlikely & \subseteq & \unlikely\tensor*[_{\mathscr{I}}]{+}{}\unlikely=\bar{\mu}\left(\proj{\psi_{2}}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(\proj{\psi_{3}}\right)\\
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi_{2}}\right)=\likely & \subseteq & \unlikely\tensor*[_{\mathscr{I}}]{+}{}\likely=\bar{\mu}\left(\proj{\psi_{3}}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(\proj{0_{\theta,\gamma}^{\perp}}\right)
\end{eqnarray*}
In order to verify $\bar{\mu}\left(\proj{\psi_{2}}\right)=\bar{\mu}\left(\proj{\psi_{3}}\right)=\unlikely$,
it is sufficient to prove that $\ket{\ps_{\theta',\gamma'}^{\perp}}\in\mathcal{B}$
implies ($0<\theta'<\frac{\pi}{2}$ and $\pi\le\gamma'<2\pi$) or
$\theta'=\frac{\pi}{2}$. Recall $\ip{\ps}{\ps_{\theta',\gamma'}^{\perp}}=0$.
Hence, if $\ket{\ps_{\theta',\gamma'}^{\perp}}\in\mathcal{B}$, we
have $\ket{\ps_{\theta',\gamma'}^{\perp}}$ parallel to $\ket{\ps}\times\ket{0_{\theta,\gamma}^{\perp}}$.
\[
\ket{\ps}\times\ket{0_{\theta,\gamma}^{\perp}}=\begin{pmatrix}1\\
1\\
0
\end{pmatrix}\times\begin{pmatrix}0\\
\rme^{\rmi\gamma}\sin\theta\\
\cos\theta
\end{pmatrix}=\begin{pmatrix}\cos\theta\\
-\cos\theta\\
\rme^{-\rmi\gamma}\sin\theta
\end{pmatrix}=\rme^{-\rmi\gamma}\begin{pmatrix}-\rme^{\rmi\left(\pi+\gamma\right)}\sin\left(\frac{\pi}{2}-\theta\right)\\
\rme^{\rmi\left(\pi+\gamma\right)}\sin\left(\frac{\pi}{2}-\theta\right)\\
\cos\left(\frac{\pi}{2}-\theta\right)
\end{pmatrix}=\rme^{-\rmi\gamma}\ket{\ps_{\frac{\pi}{2}-\theta,\pi+\gamma}^{\perp}}
\]
Because of $0\le\theta<\frac{\pi}{2}$ and $0\le\gamma<\pi$, we have
($0<\theta'=\frac{\pi}{2}-\theta<\frac{\pi}{2}$ and $\pi\le\gamma'=\pi+\gamma<2\pi$)
or $\theta'=\frac{\pi}{2}-\theta=\frac{\pi}{2}$. Similarly, when
$\ket{\psi_{1}}$ is $\ket{\ps_{\theta,\gamma}^{\perp}}$ and $\ket{\ps}\notin\mathcal{B}$,
equation (\ref{eq:non-additive-vectors}) holds. 
\item Finally, when $\ket{0}\notin\mathcal{B}$, $\ket{\ps}\notin\mathcal{B}$,
$\ket{0_{\theta,\gamma}^{\perp}}\notin\mathcal{B}$, and $\ket{\ps_{\theta,\gamma}^{\perp}}\notin\mathcal{B}$
for $0\le\theta<\frac{\pi}{2}$ and $0\le\gamma<\pi$, i.e., the ``otherwise''
case. Then, equation (\ref{eq:non-additive-vectors}) can easily be
verified. 
\begin{eqnarray*}
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi_{1}}\right)=\likely & \subseteq & \unlikely\tensor*[_{\mathscr{I}}]{+}{}\unlikely=\bar{\mu}\left(\proj{\psi_{2}}\right)\tensor*[_{\mathscr{I}}]{+}{}\bar{\mu}\left(\proj{\psi_{3}}\right)
\end{eqnarray*}
\end{enumerate}
\qed\end{example}
